Efficient spherical Convolutional Neural Networks with Healpix sampling for cosmological applications

### 2018-05-25 scnn.tex
starting to work on the data section

\subsection{Spherical convolutions}
The major challenge to define Convolution Neural Network (CNN) onto a
spherical domain is to handle a convolution suitable for this irregular domain.
They exist three approaches to address this issue.

@@

We are not the first one to leverage a graph to define a CNN on a sphere.
\cite{khasanova2017graph} use the same idea to handle omnidirectional images.
Besides the different application covered by the paper, there are two important
differences. First the sampling is different as \cite{khasanova2017graph} uses a
grid. Second, it uses a different graph CNN based on a different parameterization
of the convolution kernel.

@@

% Structure of the section
% 1) Present the goal
% 2) Explain briefly the dataset
% 3) Present the setting
% 4) Describe the result
In this section, we demonstrate the usefulness of the spherical neural network
on a discrimination problem. Our implementation is available as a small python package at \url{https://our.code.link.com}.
Furthermore, all the data to reproduce our experiments are available online.\footnote{\url{https://doi.org/10.5281/zenodo.????} \todo{correct DOI}}

\subsection{Data}
\assign{Tomek}

>>>>

\subsection{Spherical convolutions}
The major challenge to define Convolution Neural Network (CNN) onto a
spherical domain is to handle a convolution that is suitable for this irregular sampling domain.
Three approaches exist to address this issue.

@@

We are not the first one to leverage a graph to define a CNN on a sphere.
\cite{khasanova2017graph} use the same idea to handle omnidirectional images.
Besides the different application covered by the paper, there are two important
differences. First the sampling is different as \cite{khasanova2017graph} uses a
grid. Second, it uses a different graph CNN based on a different parametrisation
of the convolution kernel.

@@

% Structure of the section
% 1) Present the goal
% 2) Explain briefly the dataset
% 3) Present the setting
% 4) Describe the result
In this section, we demonstrate the performance of the spherical neural network on a discrimination problem.
Our implementation is available as a small python package\url{https://our.code.link.com}.
Furthermore, all the data to reproduce our experiments are available online.\footnote{\url{https://doi.org/10.5281/zenodo.????} \todo{correct DOI}}.

In our experiment we classify convergence maps, similar to those created with gravitational lensing technique \citep{chang2017curvedsky}.
We compare the performance of the spherical CNN to a simple benchmark algorithm.
This algorithm is based on classification of histogram of pixels in the maps.
We use an SVM to perform that task.
The benchmark method uses the information about the distribution of pixels, but does not include any spatial information.
The CNN can exploit the spatial information.
The performance is assessed by measuring the classification error on the test set.

Convergence maps represent the dimensionless distribution of over- and underdensities of mass in the universe, projected on the sky plane.
The 3D structures are projected using a geometric kernel, value of which depends on the radial distance.
In gravitational lensing, this kernel is dependent on the radial distances between the observer, the mass plane and the plane of source galaxies \citep[see][for review of gravitational lensing]{bartelman2010gravitationallensing}.
The redshifts of the sources was set to $z=1$.

We make full sky N-body simulations for two parameter set for $\Lambda \rm{CDM} $ cosmological model: model 1 ($\Omega_m=0.31, \sigma_8=0.82$) and model 2 ($\Omega_m=0.26, \sigma_8=0.91$), where $\Omega_m$ is the matter density in the universe and $\sigma_8$ is the normalisation of the matter power spectrum.
Other parameter parameters were set to $h=0.7$ (MORE PARAMS HERE - CHECK WITH RAPHAEL).
These parameters were chosen to have the same spherical harmonic power spectrum for $\ell<1000$.
That means that it is not possible to distinguish between these cosmological models using these maps if the this range of scales is used.
We found that for $\ell>1000$ the differences in power spectrum is $~5\%$.
To remove this information, we additionally smooth the spherical maps with a Gaussian kernel.

\subsection{Data}
\label{sec:data}

The simulations are created using the fast lightcone method described in \citep{sgier2018fastgeneration}.
However, we use only a single simulation box, as opposed to two used in \citep{sgier2018fastgeneration}, as we use source galaxies at lower redshift of $z=1$, instead of $z=1.5$.
As a simulator we use \pkg{L-PICOLA} \citep{howlett2015lpicola}, a fast and approximate code for N-body simulations.
For each class, we generate 30 simulations.
In the experiment we use data on a parts of the sphere, and we check the performance as a function of the area used.
We also test the performance as a function of noise level of the pixels.

% sim1: planck
% omega_m=0.31, pk_norm=0.82, h=0.7

% sim2: alternative model
% omega_m=0.26, sigma_8=0.91, h=0.7
